Towards Detecting Stalkerware Through Network Traces

### 2019-03-21 paper/main.tex
Added the filters

\documentclass[acmtog]{acmart}

\usepackage{makecell}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\renewcommand\cellalign{tl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em 
b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

Wiretapping the phones was accomplished by setting up our personal computers as 
WiFi hotspots. Since we controlled the passwords required to connect to our 
networks, and Windows alerts the user when a device connects to its mobile 
hotspot, we were able to ensure that we did not tap any traffic other than that 
of the test phones. 
\section{Results}

@@

We picked our three keywords because of xyz reasons outlined in the checkpoint

@@

In the subset of apps we studied, the words "spy," "find," and "track" appeared 
frequently in various fields of the traffic. The most common place to find them 
was in the destination domain, which we found using Wireshark's reverse DNS 
lookup tool. 12 of the 26 apps that we eventually found signatures for 
contained these spyware keywords in their domains. 
For the apps whose traffic doesn't contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.

\subsection{App Traffic}

What causes most stalkerware to send packets? Our captures and testing show 
that there are three main categories of updating tracker state: manual, timed, 
and on tracked feature. Most free apps fall into the manual update method, 
where there is a button on a website or on a phone app that will send a signal 
packet to the stalkerware app that initializes the upload. Most paid apps also 
include a timed auto-update feature, where an update will be sent every few 
minutes. A very small subset of both advertised that it receives a "real time 
notification" every time email, texts, calls, or GPS, was changed or added to. 
Our results show that none of the stalkerware apps were able to send in real 
time, all the apps advertised as real time simply updated at a more frequent 
frequency than other apps.

@@

\subsection{Network Signatures}
\label{network_signatures}

All of the above signatures are able to be filtered out through a Wireshark filter or made into a rule for a Network Intrusion Detection System such as Snort. There are many other parts of the packets that we could attempt to match signatures for, we will explain why we cannot add more to this list.
\section{Traffic Results}

The majority of stalkerware apps communicate though HTTP or HTTPS, while some 
use their own protocols. We examined the domain, SNI, User-Agent, and HTTP 
host, for each app. The words "spy," "find," and "track" appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark's reverse DNS lookup tool. 
12 of the 26 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We identified the following generalized 
signatures that can be used to identify most of our apps given nothing but 
information available on the Google Play store.

>>>>

\documentclass[acmtog]{acmart}
\usepackage{listings}

\usepackage{makecell}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{lmodern} 
\usepackage{amsmath} 
\usepackage{xcolor}  
\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	frame=single,
	breaklines=true,
}
\renewcommand\cellalign{tl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em 
b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

Wiretapping the phones was accomplished by setting up our personal computers as 
WiFi hotspots. Since we controlled the passwords required to connect to our 
networks, and Windows alerts the user when a device connects to its mobile 
hotspot, we were able to ensure that we did not tap any traffic other than that 
of the test phones. 
\section{Results}
\subsection{App Traffic}

What causes most stalkerware to send packets? Our captures and testing show 
that there are three main categories of updating tracker state: manual, timed, 
and on tracked feature. Most free apps fall into the manual update method, 
where there is a button on a website or on a phone app that will send a signal 
packet to the stalkerware app that initializes the upload. Most paid apps also 
include a timed auto-update feature, where an update will be sent every few 
minutes. A very small subset of both advertised that it receives a "real time 
notification" every time email, texts, calls, or GPS, was changed or added to. 
Our results show that none of the stalkerware apps were able to send in real 
time, all the apps advertised as real time simply updated at a more frequent 
frequency than other apps.
\subsection{Network Signatures}
\label{network_signatures}

All of the above signatures are able to be filtered out through a Wireshark filter or made into a rule for a Network Intrusion Detection System such as Snort. There are many other parts of the packets that we could attempt to match signatures for, we will explain why we cannot add more to this list.
\subsection{Traffic Results}

The majority of stalkerware apps communicate though HTTP or HTTPS, while some 
use their own protocols, such as STUN. We examined the domain, SNI, User-Agent, and HTTP 
host, for each app and tried to isolate some part that we believed would be unique enough to identify that traffic. For the majority of the apps, they have a unique domain that each app connects to, which allowed for an easy method of filtering out their traffic. Some were hosted on shared cloud servers such as AWS, and so do not have a unique domain. These, however had a phrase in user agent or in the SNI that could be used for filtering.

Initially we attempted to add on our filters with an additional check for location requests, as many stalkerware apps will contain features for tracking location. Additionally, when connected to a network a phone may use the router to find its current location. Unfortunately, while we could occasionally see a external service being used, the majority of location requests were not listed in our network captures. That being said, our resultant domain, user agent, and SNI filters managed to achieve full coverage of all stalkerware apps.

All of our filters have been collected and are listed here. These can be directly placed in Wireshark. 
\begin{lstlisting}
ip.host contains dedicated.codero.net 
|| http.host contains hellospy
|| (ip.host contains data-api.spymyfone.com && ssl.handshake.extension.data contains "my.spymyfone.com") 
|| (ip.host contains ws.spyzie.com && ssl.handshake.extension.data contains "ws.spyzie.com") 
|| ip.host contains data-api.spymyfone.com 
|| ip.host contains ws.spyzie.com 
|| ip.host contains phonedata.me 
|| ssl.handshake.extension.data contains "phonedata.me" 
|| ip.host contains apiv4.alter757.info 
|| ssl.handshake.extension.data contains "thd.cc" 
|| ip.host contains thd.cc 
|| ip.host contains mspyonline.com 
|| ip.host contains corexchange.com && ssl.handshake.extensions_server_name contains messages02.smstracker.com 
|| ip.host contains life360.helpshift.com 
|| ip.host contains antitheft.ns1.ff.avast.com 
|| (ip.host contains production-finderapps && ssl.handshake.extensions_server_name contains "api.familysafetyproduction.com") || ip.host contains production-finderapps 
|| ssl.handshake.extensions_server_name contains "celulocen.zendesk.com" 
|| ip.host contains www.zoemob.com 
|| ip.host contains unifiedlayer.com 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
|| (ip.dst_host matches "spoc-mu(|w|e)1.norton" && http.user_agent matches "(android|ios)")
|| ssl.handshake.extensions_server_name matches "family.norton" 
|| (udp && ip.host contains stun.l.google.com) 
|| http.host == relay.google.com && http.user_agent matches test 
|| http.host matches "97.logger.mobi" 
|| ssl.handshake.extensions_server_name matches "gpswox" 
|| ip.dst_host matches "gpswox" 
|| ssl.handshake.extensions_server_name matches "geozilla" 
|| ip.dst_host matches "clients.your-server.de" 
|| (ip.addr == 66.147.232.135 && tcp.port == 8080 && http.request.uri contains "signalr") 
|| ip.addr == 66.147.232.135 && tcp.port == 8080 
|| ssl.handshake.extensions_server_name matches "findmykids" 
|| ssl.handshake.extension.data contains "findmykids"  
||  ip.dst_host matches "(trackmyphones.com)" 
|| http.host matches "callsmstracker" 
|| ip.dst_host matches "onlinehome-server" 
 \end{lstlisting}

 Unfortunately while this does catch all of the stalkerware apps we have found, it does not catch most of the stalkerware apps we have not found. In this case, we need to generalize our results into a format that can be either catch-all or easily scalable without requiring a network capture.

@@

The words "spy," "find," and "track" appeared frequently in 
various fields of the traffic. The most common place to find them was in the 
destination domain, which we found using Wireshark's reverse DNS lookup tool. 
12 of the 26 apps that we eventually found signatures for contained these 
spyware keywords in their domains. We identified the following generalized 
signatures that can be used to identify most of our apps given nothing but 
information available on the Google Play store.

@@

For the apps whose traffic do not contain "spy," "find," or "track," we will 
next check for the app name and manufacturer name or substrings of these two. 
Using these indicators has a significant disadvantage - we must be certain that 
the app is stalkerware and that the manufacturer doesn't make anything except 
stalkerware. In the case of smaller apps and reskinned apps, this may actually 
be a reasonable assumption - manufacturer names are generated to fool the 
Google Play store into thinking the app is new, not to identify a real company, 
and are therefore unique by necessity. We note that the two apps we identify 
only by manufacturer are made by GPSWOX and GeoZilla. GPSWOX makes 
location-tracking apps of all kinds, some of which may be difficult to use as 
stalkerware, but GeoZilla's only apps are all designed to track a phone. Since 
only two apps were identified this way, the false positive rate should be 
investigated further.
We identified the following generalized signatures that can be used to identify 
most of our apps given nothing but information available on the Google Play 
store. The second step of our analysis will be to search for these general 
signatures in our personal phones' traffic captures. This will give a better 
idea of whether it is possible to use general signatures without generating a 
prohibitive number of false positive stalkerware detections.
\begin{lstlisting}
http.host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| ssl.handshake.extensions_server_name matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| http.user_agent matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)" 
|| ipv6.dst_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.dst_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"  
|| ipv6.src_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
|| ip.src_host matches "(spy|track|find|antitheft.ns1.ff.avast|zoemob|life360|family.norton|gpswox|geozilla)"
\end{lstlisting}

\section{Actual Traffic Testing}

@@
